Probing heavy ion collisions using quark and gluon jet substructure

### 2018-02-21 qg_ML.tex
YT updated the draft

\abstract{
We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions. We exploit jet features to highlight the differences between quark jets and gluon jets and how differently they are modified in the medium. We perform quark-gluon and proton-heavy ion jet discriminations for jets in proton-proton and heavy ion collisions. Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are examined in a multivariate analysis. In comparison, we perform the classifications using image-recognition methods by training a deep convolutional neutral network on jet images. To systematically extract jet substructure information, we develop the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales. We draw connections to the soft-drop subjet distribution and illuminate jet modifications using Lund diagrams. The quark and gluon jet samples generated from \jwpy ~are used as an example to establish our general method. We find that the quark gluon discrimination performance worsens in the \jwpy simulated heavy ion collisions. The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation. Our work suggests a systematically improvable framework for testing and comparing jet simulations and measurements in heavy ion collisions.
}

@@

The jet quenching phenomenon observed in experiments at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
and the Large Hadron Collider (LHC)~\cite{Aamodt:2010jd,Aad:2010bu,Chatrchyan:2011sx,Aamodt:2011vg,CMS:2012aa,Chatrchyan:2012nia,Abelev:2012hxa,Chatrchyan:2012gt,Aad:2012vca,
Abelev:2013kqa,Aad:2013sla,Chatrchyan:2013exa,Aad:2014bxa,Khachatryan:2014bva,
Adam:2015ewa,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha,Aad:2015wga,
Khachatryan:2016odn,Khachatryan:2016xxp,Sirunyan:2016znt,Khachatryan:2016ypw,
Sirunyan:2017isk,Sirunyan:2017oug,Sirunyan:2017lzi,Sirunyan:2017jic,Sirunyan:2017xss,Sirunyan:2017qhf}
has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions (see ~\cite{Connors:2017ptx} for a review on heavy ion jet measurements). The dramatic suppression of hadron and jet cross sections is understood using the medium-induced energy loss picture \cite{Gyulassy:1993hr,Wang:1994fx,Zakharov:1996fv,Zakharov:1997uu,Baier:1996kr,Baier:1998kq,
Gyulassy:2000er,Gyulassy:2000fs,Wiedemann:2000za,Wang:2001ifa,
Arnold:2001ba,Arnold:2001ms,Arnold:2002ja,Casalderrey-Solana:2014bpa}. In order to quantitatively extract key features of the jet-medium interaction and the underlying medium dynamics, a precise understanding of the redistribution of jet energy in heavy ion collisions is becoming essential. With the proliferation of accurate comparisons between theoretical calculations and measurements \cite{Chatrchyan:2013kwa,Khachatryan:2016tfj,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec,Khachatryan:2016erx,
Acharya:2017goa,Sirunyan:2017bsd,CMS:2017xdn}, the field of heavy ion jet physics has entered the era of precision jet substructure and jet cross section studies.

Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment. From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, we have established that both the transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructure depends strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments. It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}. With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the change of quark and gluon jet fractions caused by jet-medium interactions. This then allows one to focus on studying how differently quark and gluon jets are modified and use them as independent jet quenching probes. Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease. On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize the contributions from medium responses \cite{Tachibana:2017syd}. Quark and gluon jets will provide the opportunity for studying how differently the medium responds to different probes.

The use of quark jets and gluon jets as different hard probes of heavy ion collisions is closely related to the tagging of these two probes, which is an outstanding problem \cite{Gras:2017jty,Frye:2017yrw}. The lowest-order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons. The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets. Extra distinctive features can be included to improve the tagging by combining various jet substructure variables in a multivariate analysis, and advances in machine learning techniques have recently been incorporated which helps achieving higher tagging efficiency \cite{Komiske:2016rsd}.

Besides the conventional wisdom in jet quenching studies by directly comparing jets in proton-proton and heavy ion collisions, we will systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information. This indirect, complimentary approach studies the modification of the differences between quark jets and gluon jets which disentangles modifications of their common features. By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements. We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}. However, the method is easily applicable to other heavy ion jet simulations \cite{} and experimental data as well.

We will use three classes of methods and gain insights from comparing their performances. We combine in a multivariate analysis five physics-motivated and representative jet substructure variables which are effective in quark gluon discrimination \cite{Gallicchio:2012ez}. They include the jet mass, two radial moments including the girth, the $p_T^D$ and the pixel multiplicity. Subsets of the variables are combined and studied in a multi-layer perceptron (MLP) with two hidden layers to extract correlations among the variables. We also employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN). The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a representation of the jet energy distribution. Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}. Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method using subjets with multiple angular resolutions. At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axis which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}. The corresponding subjet expansion can be truncated at any finite order which allows us to order-by-order examine the information carried by subleading subjets. These three methods are used in discriminating quark and gluon jets, as well as discriminating jets in proton-proton and heavy ion collisions. We will see that new insights can be gained from comprehensively comparing different probes across different collision systems.

Because of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}. In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$. Within the catchment area of jets the significant background radiation needs to be subtracted before meaningful jet observable measurements can be made. Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Berta:2014eza,Komiske:2017ubm} is preferred. Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables. We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the telescoping deconstruction framework. We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent jets and illustrate the regions of phase space in the context of soft drop grooming procedure.

The rest of the paper is organized as follows. In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy ~simulation. In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as distributions of jet substructure observables. In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations. In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space separated by the soft-drop procedure. We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.

@@

To use quark and gluon jets as independent probes of heavy ion collisions, we need to first define and identify the respective probes. In this paper, we define jet flavor based on the parton from hard scattering which initiates the parton shower evolution. We use the \jwpy Monte Carlo simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}. \jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions. We use prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions. It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s. $Z$+jet) to enhance either the gluon or the quark jet fraction. This can then be directly applied to analyze other jet simulations and experimental data in the future.

In each event, jets are reconstructed using the anti-$k_{t}$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$. We impose the following kinematic cuts,
\begin{equation}
    p^{\gamma}_T > 100~{\rm GeV}, ~~~~ p^{\rm jet}_T > 50~{\rm GeV}, ~~~~ |\eta^{\rm jet}|<2, ~~~~\Delta \phi_{\gamma, {\rm jet}} > 2\pi/3\;,
\end{equation}
where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon. The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion. The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}. We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHApdf}~\cite{Whalley:2005nh}. We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons. It was shown that recoil contributions are important in \jwpy to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. The \jw simulation produces HepMC~\cite{} files which are processed using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.

@@

We use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns. In doing so one attempts to identify qualitatively different jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms. Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which show differences between them. The idea is that one can compare quark and gluon jet substructures in the same collision system and use their differences as a way to quantify the jet modification. New insights can be gained by putting jet substructures in multiple contexts. This can also help disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments. A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.

Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures \cite{}. With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful selections \cite{Gallicchio:2011xq,Gallicchio:2012ez}. On the other hand, using modern machine-learning, image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to significantly improve at high transverse momentum. This suggests that additional jet features have been identified by the deep neural network to help the tagging task. We use both methods to gain insights on how jet quenching is manifested in these two jet representations. Furthermore, we develop the telescoping deconstruction framework which represents the jet information using subjets with varying angular resolution. It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features. More details about the telescoping deconstruction algorithm and examples comparing QCD, W and top jets are provided in the Appendix.

@@

\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets. The top panels of each individual distribution show the pp ones and the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

Multivariate analyses have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study. Given a jet reconstructed using the anti-$k_{t}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
	\begin{itemize}
		\item Jet mass: $m = \sqrt{E^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
        		\item First radial moment (girth): $\sum_{i \in {\rm jet}} p^i_{T} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p^i_{T} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
        		\item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^{i}_{T}}^2}/p^{\rm jet}_{T}$.
        		\item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
	\end{itemize}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgGluonJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
\caption{Average quark (left) and gluon (right) pre-processed jet images shown in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \jw. The color scale in the z axis represents the average pixel \pt in units of GeV/c. }
\label{fig:qgjetimages}
\end{figure}

>>>>

\abstract{
We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions. We exploit jet features to highlight differences between quark jets and gluon jets and how differently they are modified in the medium. We perform quark-gluon and proton-heavy ion jet discriminations for jets in proton-proton and heavy ion collisions. Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are examined in a multivariate analysis. In comparison, we perform the classifications using image-recognition methods by training a deep convolutional neutral network on jet images. To systematically extract jet substructure information, we develop the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales. We draw connections to the soft-drop subjet distribution and illuminate jet modifications using Lund diagrams. The quark and gluon jet samples generated from \jwpy are used as an example to establish our general method. We find that the quark gluon discrimination performance worsens in the \jwpy simulated heavy ion collisions. The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy simulation. Our work suggests a systematically improvable framework for testing and comparing jet simulations and measurements in heavy ion collisions.
}

@@

The jet quenching phenomenon observed in experiments at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
and the Large Hadron Collider (LHC)~\cite{Aamodt:2010jd,Aad:2010bu,Chatrchyan:2011sx,Aamodt:2011vg,CMS:2012aa,Chatrchyan:2012nia,Abelev:2012hxa,Chatrchyan:2012gt,Aad:2012vca,
Abelev:2013kqa,Aad:2013sla,Chatrchyan:2013exa,Aad:2014bxa,Khachatryan:2014bva,
Adam:2015ewa,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha,Aad:2015wga,
Khachatryan:2016odn,Khachatryan:2016xxp,Sirunyan:2016znt,Khachatryan:2016ypw,
Sirunyan:2017isk,Sirunyan:2017oug,Sirunyan:2017lzi,Sirunyan:2017jic,Sirunyan:2017xss,Sirunyan:2017qhf}
has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions (see \cite{Connors:2017ptx} for a review on heavy ion jet measurements). The dramatic suppression of hadron and jet cross sections is understood using the medium-induced energy loss picture \cite{Gyulassy:1993hr,Wang:1994fx,Zakharov:1996fv,Zakharov:1997uu,Baier:1996kr,Baier:1998kq,
Gyulassy:2000er,Gyulassy:2000fs,Wiedemann:2000za,Wang:2001ifa,
Arnold:2001ba,Arnold:2001ms,Arnold:2002ja,Casalderrey-Solana:2014bpa}. In order to quantitatively extract key features of jet-medium interactions and the underlying medium dynamics, a precise understanding of the redistribution of jet energy in heavy ion collisions is becoming essential. With the proliferation of accurate comparisons between theoretical calculations and measurements \cite{Chatrchyan:2013kwa,Khachatryan:2016tfj,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec,Khachatryan:2016erx,
Acharya:2017goa,Sirunyan:2017bsd,CMS:2017xdn}, the field of heavy ion jet physics has entered the era of precision jet substructure and jet cross section studies.

Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment. From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, we have established that both the transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments. It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}. With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the change of quark and gluon jet fractions caused by jet-medium interactions. This then allows one to focus on studying how differently quark and gluon jets are modified and use them as independent jet quenching probes. Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease. On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}. Quark and gluon jets will provide the opportunity for studying how differently the medium responds to different probes.

The use of quark jets and gluon jets as different hard probes of heavy ion collisions is closely related to the tagging of these two probes, which is an outstanding problem \cite{Gras:2017jty,Frye:2017yrw}. The lowest-order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons. The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons results in broader spread of the radiation inside gluon jets. Extra distinctive features can be included to improve the tagging by combining various jet substructure variables in a multivariate analysis, and advances in machine learning techniques have recently been incorporated which helps achieving higher tagging efficiency \cite{Komiske:2016rsd}.

Besides the conventional wisdom in jet quenching studies by directly comparing jets in proton-proton and heavy ion collisions, we will systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information. This indirect, complimentary approach studies the modification of the differences between quark jets and gluon jets which disentangles modifications of their common features. By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements. We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}. However, the method is easily applicable to other heavy ion jet simulations \cite{Armesto:2009fj,Cao:2017zih} and experimental data as well.

We will use three classes of methods and gain insights from comparing their performances. We combine in a multivariate analysis five physics-motivated and representative jet substructure variables which are effective in quark gluon discrimination \cite{Gallicchio:2012ez}. They include the jet mass, two radial moments including the girth, the $p_T^D$ and the pixel multiplicity. Subsets of the variables are combined and studied in a multi-layer perceptron (MLP) with two hidden layers to extract correlations among the variables. We also employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN). The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a representation of the jet energy distribution. Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}. Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method using subjets with multiple angular resolutions. At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axes which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}. The corresponding subjet expansion can be truncated at any finite order which allows us to order-by-order examine the information carried by subleading subjets. These three methods are used in discriminating quark and gluon jets, as well as discriminating jets in proton-proton and heavy ion collisions. We will see that new insights can be gained from comprehensively comparing different probes across different collision systems.

Because of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}. In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$. Within the catchment area of jets the significant background radiation needs to be subtracted before meaningful jet observable measurements can be made. Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm} is preferred. Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables. We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the telescoping deconstruction framework. We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent jets and illustrate the regions of phase space in the context of soft drop grooming procedure.

The rest of the paper is organized as follows. In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy simulation. In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as the corresponding jet substructure distributions. In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations. In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space in jet formation separated in the soft-drop procedure. We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.

@@

To use quark and gluon jets as independent probes of heavy ion collisions, we need to first define and identify the respective probes. In this paper, we define jet flavor based on the parton from hard scattering which initiates the parton shower evolution. We use the \jwpy Monte Carlo simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}. \jwpy is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions. We use prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions. It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s. $Z$+jet) to enhance either the gluon or the quark jet fraction. This can then be directly applied to analyze other jet simulations and experimental data in the future.

In each event, jets are reconstructed using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$. We impose the following kinematic cuts,
\begin{equation}
    p^{\gamma}_T > 100~{\rm GeV}, ~~~ |\eta^{\rm \gamma}|<1.5, ~~~ p^{\rm jet}_T > 50~{\rm GeV}, ~~~ |\eta^{\rm jet}|<1.5, ~~~\Delta \phi_{\gamma, {\rm jet}} > 2\pi/3\;,
\end{equation}
where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon. The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion. The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}. We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHAPDF}~\cite{Whalley:2005nh}. We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons. It was shown that recoil contributions are important in \jwpy to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. In order to remove the thermal component of recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. The \jw simulation produces HepMC~\cite{} files which are processed using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.

@@

We use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns. In doing so one attempts to identify qualitatively different jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms. Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which show differences between them. The idea is that one can compare quark and gluon jet substructures in the same collision system and use their differences as a way to quantify the jet modification. New insights can be gained by putting jet substructures in multiple contexts. This can also help disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments. A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Figure~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.

Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures \cite{}. With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful selections \cite{Gallicchio:2011xq,Gallicchio:2012ez}. On the other hand, using modern machine-learning, image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to significantly improve at high transverse momentum. This suggests that additional jet features have been identified by the deep neural network to help the tagging task. We use both methods to gain insights on how jet quenching is manifested in these two jet representations. Furthermore, we develop the telescoping deconstruction framework which represents the jet information using subjets with varying angular resolution. It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features. More details about the telescoping deconstruction algorithm and examples comparing QCD, $W$ and top jets are provided in the Appendix. We will show and discuss the classification performances of all the methods in the next section.

@@

\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets. The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

Multivariate analyses have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study. Given a jet reconstructed using the anti-$k_{T}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
	\begin{itemize}
		\item Jet mass: $m = \sqrt{(\sum_{i\in {\rm jet}} E_i)^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
        \item First radial moment (girth): $\sum_{i \in {\rm jet}} p_{T_i} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p_{T_i} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
        \item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^2_{T_i}}}/p^{\rm jet}_{T}$.
        \item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
	\end{itemize}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgGluonJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
\caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}. The color scale represents the average transverse momenta deposited in each pixel in units of GeV/c. }
\label{fig:qgjetimages}
\end{figure}
